On Sample Size Needed for Block Bootstrap Confidence Intervals to Have Desired Coverage Rates

### 2023-08-30 Manuscript/manuscript.tex
Responded to Comments

\jy{What was $B$?}

Three target parameters of $X_t$ were considered:
1) $\mu = 0$, the mean of $X_t$;
2) $\sigma_x = 1$, the standard deviation of $X_t$; and
3) $\phi$, the lag-1 autocorrelation coefficient.
To investigate the effect of sample size~$n$, we considered an array of values
$n \in \{100, 200, 400, 800, 1600, 3200\}$. In each configuration, we generated 
10000 replicates. For each replicate, we constructed six 95\% block bootstrap 
CIs for each parameter as described in the last section. We can estimate $\mu$, 
$\sigma_x$, and $\phi$ by computing the sample mean, sample standard deviation, 
and lag-1 autocorrelation, respectively, of each bootstrap sample. Then we can 
construct intervals for each parameter using the appropriate procedures 
described in \textit{Block Bootstrap CIs}. 
Then we estimated their actual coverage 
rates along with their 95\% confidence intervals from the 10000 replicates. The 
block bootstrap sampling step was done with function \texttt{tsboot} from R 
package \textsl{boot},\citep{boot} with block size $\lceil n / l \rceil$. This 
function by default is an implementation of moving block bootstrap as described 
in the previous section, meaning that that blocks are allowed to wrap around, 
and $l = \lceil n^{1/3} \rceil$.

@@

The outcomes of the $\phi$ estimation raise a natural question about the 
lackluster performance of certain methodologies. To delve into this inquiry, a
set of 50 CIs was generated for each of the percentile, BC, and recentered
percentile  approaches for samples of $n \in \{800, 1600\}$. Illustrated in 
\textbf{Figure~\ref{fig:npi}}, it becomes evident that the percentile-based CIs 
exhibit a notable bias, predominantly manifesting as a substantial 
underestimation of $\phi$ with point estimator $\bar\theta_n^B$, that is, the 
average of $B$ bootstrap point estimates. As the sample size increases from 800 
to 1600, this bias does not vanish while the uncertainty reduces, which explains 
why the coverage rates deteriorate. The bias in $\bar\theta_n^B$ also appears to 
invalidate the bias-correction in the BC bootstrap, leading to the poor 
performance of the BC intervals. The BCA intervals have the same problem as the 
BC intervals in the bias-correction step. The root of the issue appears to be 
that the autocorrelation in the block bootstrap samples is somehow smaller 
compared to that in the original sample. On the other hand, the original point 
estimate $\hat\theta_n$ is asymptotically unbiased. Since the width is based on 
the uncertainty in the bootstrap point estimates $\hat\theta_n^{(b)}$, 
$b = 1, \ldots, B$, the percentile CIs recentered at the original point estimate
$\hat\theta_n$ provide desired coverage.

@@

An additional simulation study in which block bootstrap is used to estimate
\jy{not an AR(1) process...}
parameters of an AR(1) process with an underlying exponential marginal 
distribution is discussed in the Appendix. Overall, the results with an 
underlying marginal exponential distribution are very similar to the results 
with an underlying normal distribution discussed above.

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_exp_mu}
  \caption{Empirical coverage rates of different 95\% block bootstrap CIs for
    the marginal mean $\mu$ of a marginally transformed AR(1) process with an 
    underlying exponential distribution ($\lambda = 1$) with AR 
    coefficient
    $\phi \in \{-0.4, 0.2, 0, 0.2, 0.4\}$ and series length
    $n \in \{100, 200, 400, 800, 1600, 3200\}$ with
    an exponential distribution with $\lambda = 1$ based on 10,000 replicates. 
    The error bars represent 95\% CIs of the real coverage rates.}
  \label{fig:exp_mu}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_exp_sigma}
  \caption{Empirical coverage rates of different 95\% block bootstrap CIs for
    the marginal standard deviation $\sigma_x$ of a marginally transformed AR(1) 
    process with an 
    underlying exponential distribution ($\lambda = 1$) with AR 
    coefficient $\phi \in \{-0.4, 0.2, 0, 0.2, 0.4\}$ and series length 
    $n \in \{100, 200, 400, 800, 1600, 3200\}$ transformed to
    an exponential distribution with $\lambda = 1$ based on 10,000 replicates. 
    The error bars represent 95\% CIs of the real coverage rates.}
  \label{fig:exp_sigma}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_exp_phi}
  \caption{Empirical coverage rates of different 95\% block bootstrap CIs for 
    the first-order autocorrelation coefficient $\phi$ of a marginally 
    transformed AR(1) process with 
    an 
    underlying exponential distribution ($\lambda = 1$) with
    $\phi \in \{-0.4, 0.2, 0, 0.2, 0.4\}$ and series length
    $n \in \{100, 200, 400, 800, 1600, 3200\}$ transformed to
    an exponential distribution with $\lambda = 1$ based on 10,000 replicates. 
    The error bars represent 95\% CIs of the real coverage rates.}
  \label{fig:exp_phi}
\end{figure}

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/exp_pho_intervals}
  \caption{50 replicate percentile, BC, and recentered percentile CIs for 
    samples of $n \in \{800, 1600\}$ for an AR(1) process with $\phi = 0.4$
    \jy{not accurate}
    transformed to an exponential distribution iwth $\lambda = 1$. For 
    each replicate, the lower and upper bounds of the CIs are displayed, as well 
    as $\hat\theta_n$ (blue circle) and $\bar\theta_n^{B}$ (red cross).}
  \label{fig:eri}
\end{figure}

It seems that range of sample sizes needed 
to cover parameters of an exponentially distributed AR(1) process is generally 
higher than that needed to cover parameters of a normally
distributed AR(1) process. Despite this, the trends and patterns discussed 
previously with respect
to the performance of different methods and with different parameters are the 
same.
For instance, Student's $t$ CIs still appear to yield higher coverage rates than 
other 
methods. Performance is still generally better when temporal dependence is 
negative rather than positive. Most importantly, percentile, BC, and BCA CIs
still deteriorate in coverage of $\phi$ as sample size increases. As exhibited 
in Figure~\ref{fig:eri}, percentile and BC CIs still have the same bias
issue. Overall, the results with an underlying marginal exponential distribution 
are very similar to the results with an underlying normal distribution from the
main study.

\jy{comment on that the recentered percentile CI for autocorrelation still works
  as the point estimator is unbiased.}

>>>>

\jy{What was $B$?} \mc{addressed}

Three target parameters of $X_t$ were considered:
1) $\mu = 0$, the mean of $X_t$;
2) $\sigma_x = 1$, the standard deviation of $X_t$; and
3) $\phi$, the lag-1 autocorrelation coefficient.
To investigate the effect of sample size~$n$, we considered an array of values
$n \in \{100, 200, 400, 800, 1600, 3200\}$. In each configuration, we generated 
10000 replicates. For each replicate, we constructed six 95\% block bootstrap 
CIs for each parameter as described in the last section with $B = 1000$. We can 
estimate $\mu$, 
$\sigma_x$, and $\phi$ by computing the sample mean, sample standard deviation, 
and lag-1 autocorrelation, respectively, of each bootstrap sample. Then we can 
construct intervals for each parameter using the appropriate procedures 
described in \textit{Block Bootstrap CIs}. 
Then we estimated their actual coverage 
rates along with their 95\% confidence intervals from the 10000 replicates. The 
block bootstrap sampling step was done with function \texttt{tsboot} from R 
package \textsl{boot},\citep{boot} with block size $\lceil n / l \rceil$. This 
function by default is an implementation of moving block bootstrap as described 
in the previous section, meaning that that blocks are allowed to wrap around, 
and $l = \lceil n^{1/3} \rceil$.

@@

The outcomes of the $\phi$ estimation raise a natural question about the 
lackluster performance of certain methodologies. To delve into this inquiry, a
set of 50 CIs was generated for each of the percentile, BC, and recentered
percentile  approaches for samples of $n \in \{800, 1600\}$. Illustrated in 
\textbf{Figure~\ref{fig:npi}}, it becomes evident that the percentile-based CIs 
exhibit a notable bias, predominantly manifesting as a substantial 
underestimation of $\phi$ with point estimator $\bar\theta_n^B$, that is, the 
average of $B$ bootstrap point estimates. As the sample size increases from 800 
to 1600, this bias does not vanish while the uncertainty reduces, which explains 
why the coverage rates deteriorate. The bias in $\bar\theta_n^B$ also appears to 
invalidate the bias-correction in the BC bootstrap, leading to the poor 
performance of the BC intervals. The BCA intervals have the same problem as the 
BC intervals in the bias-correction step. The root of the issue appears to be 
that the autocorrelation in the block bootstrap samples is somehow smaller 
compared to that in the original sample. On the other hand, the original point 
estimator $\hat\theta_n$ is asymptotically unbiased. Since the width is based on 
the uncertainty in the bootstrap point estimates $\hat\theta_n^{(b)}$, 
$b = 1, \ldots, B$, the percentile CIs recentered at the original point estimate
$\hat\theta_n$ provide desired coverage.

@@

An additional simulation study in which block bootstrap is used to estimate
\jy{not an AR(1) process...} \mc{addressed}
parameters of a marginally transformed AR(1) process with an underlying 
exponential distribution is discussed in the Appendix.

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_exp_mu}
  \caption{Empirical coverage rates of different 95\% block bootstrap CIs for
    the marginal mean $\mu$ of a marginally transformed AR(1) process with AR 
    coefficient
    $\phi \in \{-0.4, 0.2, 0, 0.2, 0.4\}$ with an 
    underlying exponential distribution ($\lambda = 1$) and series length
    $n \in \{100, 200, 400, 800, 1600, 3200\}$ with
    an exponential distribution with $\lambda = 1$ based on 10,000 replicates. 
    The error bars represent 95\% CIs of the real coverage rates.}
  \label{fig:exp_mu}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_exp_sigma}
  \caption{Empirical coverage rates of different 95\% block bootstrap CIs for
    the marginal standard deviation $\sigma_x$ of a marginally transformed AR(1) 
    process with AR 
    coefficient $\phi \in \{-0.4, 0.2, 0, 0.2, 0.4\}$ with an 
    underlying exponential distribution ($\lambda = 1$) and series length 
    $n \in \{100, 200, 400, 800, 1600, 3200\}$ based on 10,000 replicates. 
    The error bars represent 95\% CIs of the real coverage rates.}
  \label{fig:exp_sigma}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_exp_phi}
  \caption{Empirical coverage rates of different 95\% block bootstrap CIs for 
    the first-order autocorrelation coefficient $\phi$ of a marginally 
    transformed AR(1) process with
    $\phi \in \{-0.4, 0.2, 0, 0.2, 0.4\}$ with 
    an 
    underlying exponential distribution ($\lambda = 1$) and series length
    $n \in \{100, 200, 400, 800, 1600, 3200\}$ based on 10,000 replicates. 
    The error bars represent 95\% CIs of the real coverage rates.}
  \label{fig:exp_phi}
\end{figure}

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/exp_pho_intervals}
  \caption{50 replicate percentile, BC, and recentered percentile CIs for 
    samples of $n \in \{800, 1600\}$ for a marginally transformed AR(1) process 
    with $\phi = 0.4$
    \jy{not accurate} \mc{is this better}
    with an underlying exponential distribution ($\lambda = 1$). For 
    each replicate, the lower and upper bounds of the CIs are displayed, as well 
    as $\hat\theta_n$ (blue circle) and $\bar\theta_n^{B}$ (red cross).}
  \label{fig:eri}
\end{figure}

It seems that range of sample sizes needed 
to cover parameters of an exponentially distributed AR(1) process is generally 
higher than that needed to cover parameters of a normally
distributed AR(1) process. Despite this, the trends and patterns discussed 
previously with respect
to the performance of different methods and with different parameters are the 
same.
For instance, Student's $t$ CIs still appear to yield higher coverage rates than 
other 
methods. Performance is still generally better when temporal dependence is 
negative rather than positive. Most importantly, percentile, BC, and BCA CIs
still deteriorate in coverage of $\phi$ as sample size increases. As exhibited 
in Figure~\ref{fig:eri}, percentile and BC CIs still suffer from the same bias
issue, whereas recentered percentile CI still work when estimating $\phi$ 
because the 
original point estimator is unbiased. Overall, the results with an underlying 
marginal 
exponential distribution 
are very similar to the results with an underlying normal distribution from the
main study.

\jy{comment on that the recentered percentile CI for autocorrelation still works
  as the point estimator is unbiased.} \mc{addressed}
